---
title: Analyzing Soy Field Trial
author: Omer Perach
date: '2020-06-19'
slug: how-you-can-build-a-dashboard-to-you-field-trials-data
categories: []
tags:
  - Data Visualization
  - Plant field trials
  - Plant Breeding
---

You are a plant breeder, you in the end of collecting all the data for a long and exhaustive four year project with the goal to find the crop lines with the best traits for the customers.

Let's take a data set from the `agridat` package:
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(agridat)
library(tidyverse)
library(ggplot2)
data("australia.soybean")
```

The data is a field trial on 58 varieties of soybean, grown in four locations across two years in Australia. 
Now we look on the six first rows of the data:
```{r}
head(australia.soybean)
```

We can also summarize the data set:
```{r echo=TRUE}
summary(australia.soybean)
```

##### As we can see from the summary:
1. env variable is just a representation of place and year, first letter of location and the last two character of the year.
2. Location- because there are 58 varieties multiple by two years 1970 and 1971 we have 116 observations for each location.
3. Lodging - lodging is a level of the soy stem to bend, if the stem is bending too much, it can have effect on light penetration thus on yield.I cannot understand the Lodging units. 
4. Size - seed size in millimeter ,w e can see very high range from 4 to 23.
5. protein - percentage, probably from wet matter. 
6. Oil - percentage probably from wet matter.
7. Gene:
    + Lines 1-40 Local sections for Australia
    + Line 41 Avoyelles (CPI 15939) Tanzania
    + Line 42 Hernon 49 (CPI 15948) Tanzania
    + Line 43 Mamloxi (CPI 172) Nigeria
    + Line 44 Dorman USA 
    + Line 45 Hampton USA
    + Line 46 Hill USA
    + Line 47 Jackson USA
    + Line 48 Leslie USA
    + Line 49 Semstar Australia
    + Line 50 Wills USA
    + Line 51 C26673 Morocco
    + Line 52 C26671 Morocco
    + Line 53 Bragg USA
    + Line 54 Delmar USA
    + Line 55 Lee USA
    + Line 56 Hood USA
    + Line 57 Ogden USA
    + Line 58 Wayne USA


Let's inquire each continuous variable individually for 1970 per strain:
```{r echo=TRUE}
df_1970<-australia.soybean%>%filter(year==1970)
label<-seq(1,58)
label
df_1970%>%ggplot(aes(gen,size))+geom_jitter(aes(color=loc))+coord_flip()+scale_x_discrete(label=label)
```

Now we will look on the protein percentage: 
```{r echo=TRUE}
df_1970%>%ggplot(aes(gen,protein))+geom_jitter(aes(color=loc))+coord_flip()+scale_x_discrete(label=label)
```

Height:
```{r echo=TRUE}
df_1970%>%ggplot(aes(gen,height))+geom_jitter(aes(color=loc))+coord_flip()+scale_x_discrete(label=label)
```
Lodging:
```{r echo=TRUE}
df_1970%>%ggplot(aes(gen,lodging))+geom_jitter(aes(color=loc))+coord_flip()+scale_x_discrete(label=label)
```

yield:
```{r echo=TRUE}
df_1970%>%ggplot(aes(gen,yield))+geom_jitter(aes(color=loc))+coord_flip()+scale_x_discrete(label=label)
```
Oil:
```{r echo=TRUE}
df_1970%>%ggplot(aes(gen,oil))+geom_jitter(aes(color=loc))+coord_flip()+scale_x_discrete(label=label)
```

##### Link to the dashboard I built for this data
[Dashboard](https://omerperach.shinyapps.io/agridata/)