---
title: How can you build a dashboard to you field trials data
author: Omer Perach
date: '2020-06-19'
slug: how-you-can-build-a-dashboard-to-you-field-trials-data
categories: []
tags:
  - Data Visualization
  - Plant field trials
  - Plant Breeding
description: ''
topics: []
---

You are a plant breeder, you in the end of collecting all the data for a long and exhaustive four year project with the goal to find the crop lines with the best traits for the customers.

Let's take a data set from the `agridat` package:
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(agridat)
data("australia.soybean")
```

The data is a field trial on 58 varieties of soybean, grown in four locations across two years in Australia. 
Now we look on the six first rows of the data:
```{r}
head(australia.soybean)
```

We can also summarize the data set:
```{r echo=TRUE}
summary(australia.soybean)
```

##### As we can see from the summary:
1. env variable is just a representation of place and year, first letter of location and the last two character of the year.  