---
title: 'RFM Model '
author: Omer Perach
date: '2020-06-15'
slug: rfm-model
categories: []
tags: []
description: ''
topics: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RFM analysis  - Introduction
RFM helps to identify customers who are more likely to respond to promotions by segmenting them into various categories.
We will use the **rfm** package. We will read the csv file *RFM_data.csv* and calculate RFM scores for each customer.
The data include 39999 observations of:
1. Customer I.D
2. Recencency Days
3. Number of Orders
4. Total Revenue
5. Most Recent Visit *Will not be used*
Four bins will be used for **recency**, **frequency** and **monetary**. In addition, we will present some graphs to see the recency, frequency and monetary scores for the customers.
In the end, we will divide our customers into 11 categories of customers.

##### RFM Table

In the following steps we used the following [reference](https://rfm.rsquaredacademy.com/articles/rfm-customer-level-data.html).

```{r message=FALSE, warning=FALSE, include=FALSE}
library(rfm)
library(lubridate)
library(dplyr)
library(tidyverse)
library(ggplot2)
```
```{r echo=TRUE}
df<-read.csv("RFM_data.csv",header = T)
head(df)
nrow(df)
df$most_recent_visit<-as.Date(df$most_recent_visit, format="%d/%m/%y")
head(df)
analysis_date <- lubridate::as_date("2020-03-21", tz = "UTC")
rfm_score<-rfm_table_customer(data = df,customer_id = customer_id,n_transactions =number_of_orders,recency_days = recency_days,total_revenue = revenue,analysis_date = analysis_date,recency_bins = 4,frequency_bins =4, monetary_bins = 4)
rfm_score[1:10,]
```
##### Exploratory Data Analysis of RFM Table
```{r}
rfm_heatmap(rfm_score) #combination of recency vs frequency colored by monetary
rfm_bar_chart(rfm_score) #same concept but in bar graph
rfm_order_dist(rfm_score) # Number of customers per orders
```
##### Continuous graphs 
```{r}
rfm_fm_plot(rfm_score) #Frequency vs monetary
rfm_rm_plot(rfm_score) #Recency ve monetary
rfm_rf_plot(rfm_score) #Recency vs frequency
```

##### Segments
In the following section we are following this [reference] (https://medium.com/@triimamwicaksono_47213/customer-segmentation-and-strategy-using-rfm-analysis-in-rstudio-be79118c8235) in order to classify our customers into the below segment defenitions:  
1. Champions
2. Loyal Customers
3. Potential Loyalist
4. Recent Customers
5. Promising
6. Customers Needing Attention
7. About to Sleep
8. At Risk
9. Can't Lose Them
10. Hibernating
11.Lost

```{r echo=TRUE}
champions<- c(444)
loyal_customers <- c(334, 342, 343, 344, 433, 434, 443)
potential_loyalist <-c(332,333,341,412,413,414,431,432,441,442,421,422,423,424)
recent_customers <- c(411)
promising <- c(311, 312, 313, 331)
needing_attention <- c(212,213,214,231,232,233,241,314,321,322,323,324)
about_to_sleep <- c(211)
at_risk <- c(112,113,114,131,132,133,142,124,123,122,121,224,223,222,221)
cant_lose <- c(134,143,144,234,242,243,244)
hibernating <- c(141)
lost <- c(111)
rfm_scores<-as.vector(rfm_score$rfm$rfm_score)
rfm_scores[which(rfm_score$rfm$rfm_score %in% champions)]="Champions"
rfm_scores[which(rfm_scores %in% potential_loyalist)] = "Potential Loyalist"
rfm_scores[which(rfm_scores %in% loyal_customers)] = "Loyal Customers"
rfm_scores[which(rfm_scores %in% recent_customers)] = "Recent Customers"
rfm_scores[which(rfm_scores %in% promising)] = "Promising"
rfm_scores[which(rfm_scores %in% needing_attention)] = "Customer Needing Attention"
rfm_scores[which(rfm_scores %in% about_to_sleep)] = "About to Sleep"
rfm_scores[which(rfm_scores %in% at_risk)] = "At Risk"
rfm_scores[which(rfm_scores %in% cant_lose)] = "Can't Lose Them"
rfm_scores[which(rfm_scores %in% hibernating)] = "Hibernating"
rfm_scores[which(rfm_scores %in% lost)] = "Lost"
customer_sement<-data.frame(cus_seg=rfm_scores)
customer_sement%>%count(cus_seg)%>%arrange(desc(n))%>%rename(cus_seg = cus_seg, Count = n)
ggplot(data = customer_sement) + aes(x = cus_seg, fill = cus_seg)+ geom_bar() + labs(title = "Customer Segmentation", x = "Segment", y = "Total Customer") + coord_flip()+ theme_minimal()
```
Finally I think it will be best to add the classification into the **df** data frame we start from:
```{r}
df<-df%>%mutate(Segment_name=customer_sement$cus_seg)
head(df)
```