---
title: How can you build a dashboard to you field trials data
author: Omer Perach
date: '2020-06-19'
slug: how-you-can-build-a-dashboard-to-you-field-trials-data
categories: []
tags:
  - Data Visualization
  - Plant field trials
  - Plant Breeding
description: ''
topics: []
---

You are a plant breeder, you in the end of collecting all the data for a long and exhaustive four year project with the goal to find the crop lines with the best traits for the customers.

Let's take a data set from the `agridat` package:
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(agridat)
library(tidyverse)
library(ggplot2)
data("australia.soybean")
```

The data is a field trial on 58 varieties of soybean, grown in four locations across two years in Australia. 
Now we look on the six first rows of the data:
```{r}
head(australia.soybean)
```

We can also summarize the data set:
```{r echo=TRUE}
summary(australia.soybean)
```

##### As we can see from the summary:
1. env variable is just a representation of place and year, first letter of location and the last two character of the year.
2. Location- because there are 58 varieties multipe by two years 1970 and 1971 we have 116 observations for each location.
3. Lodging - lodging is a level of the soy stem to bend, if the stem is bending too much, it can have effect on light penetration thus on yield.I cannot understand the Lodging units. 
4. Size - seed size in milimieter ,w e can see very high ragnge from 4 to 23.
5. protein - percentage, probably from wet matter. 
6. Oil - percentage probably from wet matter.

Let's inquire each continuous variable individually for 1970 per strain:
```{r echo=TRUE}
df_1970<-australia.soybean%>%filter(year==1970)
label<-seq(1,58)
label
df_1970%>%ggplot(aes(gen,size))+geom_jitter(aes(color=loc))+coord_flip()+scale_x_discrete(label=label)
```
