---
title: "GenoSee Business Model"
author: "Omer Perach"
date: '2020-07-30'
output: pdf_document
description: ''
tags:
- Finance
- SaaS
- Plant Breeding
categories: []
topics: []
---

## Executive Summary
Wordlwide climate change and decrease in arable lands and lands under permanent crops constitute a major
concern for future food safety. Private and goverment based breeding programs working constantly to develop
elite crops in order to meet the future demand to feed 10 billion people by 2050. Under these constraints, to meet the rising demand for food, agricultural production must increase by an estimated 50% without greatly increasing water usage or expanding the total land area dedicated to agriculture.
The era of data enable us to make processes highly effieciet for companies and organisations. This is also true for the plant breeding industry, huge amount of data is recorded and collected - genotype, phenotype, environmental etc. Before inferring and analyzing the data, it should be transformed into findable, available, identifiable, reusable database. 
GenoSee is a one stop shop for all the breeding program associates which let them focus in achieving their goal - developing elite crops  - while letting GenoSee handle the data related work. 
GenoSee estimate that it will reach 50 paying breeding programs worldwide during the next four years. We are asking for **$1M** dollars for the first phase of development and **$500K** for the second phase of devlopment. 
Following our assumptions we calculte a potential positive **NPV**.

## Assumptions

1. Thousands of open plant breeding programs worldwide.
2. Phase 1 - one year to MVP and reaching first customers.
3. Phase 2 - one year for improvement and growth.
4. We will look on the next 7 years.
5. Tax rate 38%.
6. Discount rate 12%.
7. Depreciation 4 years straight line.
8. During the next four years we will reach 50 breeding programs (to our opinion it is a conservative aproach).
9. GenoSee will charge 15000 dollars for building the database abd additional 300 dollars per user monthly.
10. GenoSee estimating average of 10 users per breeding program. 

## CAPEX Block
We will start by creating a depreciation matrix for the capital investment of the company. We will assume 4
years straight line depreciation that will taken into account only until it can be applied against taxable profit .

```{r echo=TRUE}
kHorizon<-7 # The model Horizon
year<-1:kHorizon # A sequence to represent the horizon years
kTaxRate<-38 # Tax percentage
kDiscountrate<-12 # Discount rate percentage per year for NPV calculations.
kDeprPer<-4 #year, the deprciation sheduale for the capital
p1.dur<-1 #Duration phase one in year
p2.dur<-1 #Duration phase 2 in year
p1.capex<-1000000 #Investment for phase 1.
p2.capex<-500000 #Investment phase 2.
maint.capex<-0 # Additional parameter nit used.
time.to.peak.sales<-3 #Time to reach 100 breedings programs.
mkt.demand<-50 # Market demand after 4 years.
phase <- (year <= p1.dur) * 1 +
(year > p1.dur & year <= (p1.dur + p2.dur)) * 2 +
(year > (p1.dur + p2.dur)) * 3
capex<-(phase==1)*p1.capex/p1.dur+(phase==2)*p2.capex/p2.dur+(phase==3)*maint.capex
depr.matrix <-
t(sapply(year, function(y)
ifelse(y <= p1.dur & year > 0,0,ifelse(
y == (p1.dur + 1) & year < y + kDeprPer & year >= y,
p1.capex / kDeprPer,
ifelse((year >= y) & (year < (y + kDeprPer)),
capex[y - 1] / kDeprPer, 0)
)
)
)
)
depr.matrix
depr<-colSums(depr.matrix)
plot(year,depr/1e3,xlab="Year",ylab="Depreciation[$000]",type ="b" )
```

## Sales and Revenue Block

We assume that it will take us 3 years after end of phase 1 to reach 50 breeding programs worldwide. We
will price (assumption) the software to be 15000 dollars per year and additional subscription price per user of
300 dollars per month. We estimate that each program will have 5 users by average.

```{r echo=TRUE}
mkt.adoption<-pmin(cumsum(phase>1)/time.to.peak.sales,1)
price.per.software.year<-15000
price.per.user<-300
average.users.per.program<-10
month.per.year<-12
user.arr<-price.per.user*average.users.per.program*month.per.year
price<-price.per.software.year+user.arr
sales<-mkt.adoption*mkt.demand
revenue<-sales*price
plot(year, revenue/1000000,
xlab = "Year",
ylab = "Sales [000,000 $]",
type = "b")
```

## OPEX Block
The OPEX include R&D (research and development) and G&S (general sales and administrative) and is estimated to be 20% of the revenue. The operational cost will start only after phase 1.
```{r echo=TRUE}
opex<-revenue*0.2
plot(year, opex/1000,
xlab = "Year",
ylab = "Sales [000$]",
type = "b")
```

## Pro Forma Table
+ Gross profit = revenue-OPEX
+ Operating profit before taxes = Gross profit - OPEX - depreciation
+ Operating profit after tax = operating profit before tax-tax
+ Cash flow = operating profit after tax + depreciation - CAPEX

```{r}
gross.profit<-revenue-opex
op.profit.before.tax<-gross.profit-opex-depr
tax<-op.profit.before.tax*kTaxRate/100
op.profit.after.tax<-op.profit.before.tax-tax
cash.flow<-op.profit.after.tax+depr-capex
cum.cah.flow<-cumsum(cash.flow)
```

## Cumulative Cash Flow
```{r}
plot(year, cum.cah.flow/1000,xlab = "Year",ylab = "Cash Flow [000$]", type = "b")
```

## Net Present Value
```{r}
discount.factors <- 1/(1 + kDiscountrate / 100) ^ year
discounted.cash.flow <- cash.flow * discount.factors
plot(year,
discounted.cash.flow,
xlab = "Year",
ylab = "Discounted Cash Flow [$000]",
type = "b")
npv <- sum(discounted.cash.flow)
```

**Net Present Value is** `r npv`

## Pro Forma Table

```{r}
pro.forma.vars <- array( c(sales, revenue, gross.profit,
-opex, -depr, op.profit.before.tax, -tax, op.profit.after.tax, depr, -capex, cash.flow ), dim = c(kHorizon, 11))
pro.forma <- data.frame(pro.forma.vars)
pro.forma.headers <- c("Sales [breeding programs]", "Revenue","Gross Profit",
"OPEX", "-Depreciation", "Operating Profit Before Tax", "Tax", "Operating Profit After Tax", "+Depreciation", "CAPEX",
"Cash Flow")
colnames(pro.forma) <- pro.forma.headers
rownames(pro.forma) <- year
pro.forma = t(pro.forma)
pro.forma
```

## References
The business model in this post based on a book by Robert D. Brown III *Business Case Analysis with R**