---
title: "Looking on your home from space"
author: "Omer Perach"
date: '2021-01-14'
slug: looking-on-your-home-from-space
categories: []
tags:
- Remote Sensing
- Raster Image Processing
- Vegetative Index
description: ''
topics: []
---

# The incentive 

If you are working as a farmer and you know that you can monitor your fields for free. Would you not do it?
**This post is to teach you how to download data from a satellite called Sentinel and to process it for free, not paying anyone to do it for you or you doing it but with the aid of a very expansive software**

# How to dowanload the images

There are bunch of R packages that can help you to download images from many free satellites such as:  

1. RGIStools
2. getspatialdata
3. sen2r (only for Sentinel)

I used sen2r to download the images.
what I love in sent2 is the shiny GUI interface they build in order to help you download the data. I will suggest you to read this [link](https://www.sciencedirect.com/science/article/pii/S0098300419304893?via%3Dihub) - it explaining the basic information for knowing how to use it.
I will not go into it in this post but if you would like me to elaborate on it, please contact me.\
Let's assume that you read the instructions and you already download the pictures into your compute.

# Reading the pictures
We will use the *raster* package and basically each picture is a multylayer of all the bends, sen2R 
```{r Reading rasters, echo=TRUE, message=FALSE, warning=FALSE}
library(raster)
library(terra)
library(rasterVis)
pic_21<-rast("S2A2A_20201221_121_sen2r_BOA_10.tif")
pic_26<-rast("S2B2A_20201226_121_sen2r_BOA_10.tif")
pic_31<-rast("S2A2A_20201231_121_sen2r_BOA_10.tif")
```

## Band ordered when downloaded with the package:
1. - 442.7 nm (Coastal aerosol)
2. - 492.4 nm (blue)
3. - 559.8 nm (Green)
4. - 664.6 nm (Red)
5. - 704.1 nm (Vegetation Red Edge)
6. - 740.5 nm (Vegetation Red Edge)
7. - 782.8 nm (Vegetation Red Edge)
8. - 832.8 nm (below 20m rsolution)/864.7 nm (above 20m resolution) (Nir)/ (Vegetation Red Edge)
9. - 945.1 nm (Water vapour)
10. - 1613.7 nm (SWIR)
11. - 2202.4 nm (SWIR)

Let's print RGB pictures just to see and placed on a base map just to see where we are.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(mapview)
pic_21_raster<-stack(pic_21)
mapview::viewRGB(pic_21_raster,r=3,b=1,g=2)
```

Great, the fields of my childhood :)

Now let's try to create NDVI raster and plot it 

```{r echo=TRUE, message=FALSE, warning=FALSE}
ndvi_21<-(pic_21_raster[[5]]-pic_21_raster[[4]])/(pic_21_raster[[5]]+pic_21_raster[[4]])
mapview::mapview(ndvi_21)
```

Now, it is not very common to see NDVI as above, it is more convenient to use other palette,most of the times between red and green.
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(RColorBrewer)
ndvi_palette<-brewer.pal(9,"RdYlGn")
mapview::mapview(ndvi_21,col.region=ndvi_palette)
```


