---
title: Looking on your home from space
author: Omer Perach
date: '2021-01-14'
slug: looking-on-your-home-from-space
categories: []
tags:
  - Remote Sensing
  - Raster Image Processing
  - Vegetative Index
description: ''
topics: []
---

# The incentive 
If you do not know it, there are satellites in the sky:)\
I am joking, most of us are aware about it, but I don't think that most of us are aware that satellites traveling in the sky and taking pictures that we could download and process.\
**This post is to teach you how to download data from a satellite called Sentinel and to process it**

# How to dowanload the images
There are bunch of R packages that can help you to download images from many free satellites such as:\
1. RGIStools\
2. getspatialdata\
3. sen2r (only for Sentinel)\
I used sen2r to download the images.
what I love in sent2 is the shiny GUI interface they build in order to help you download the data. I will suggest you to read this [link](https://www.sciencedirect.com/science/article/pii/S0098300419304893?via%3Dihub) - it explaining the basic information for knowing how to use it.
I will not go into it in this post but if you would like me to elaborate on it, please contact me.\
Let's assume that you read the instructions and you already download the pictures into your compute.

# Reading the pictures
We will use the *raster* package and basically each picture is a multylayer of all the bends, sen2R 
```{r}
library(raster)
library(rasterVis)
pic_21<-stack("S2A2A_20201221_121_sen2r_BOA_10.tif")
pic_26<-stack("S2B2A_20201226_121_sen2r_BOA_10.tif")
pic_31<-stack("S2A2A_20201231_121_sen2r_BOA_10.tif")
```

## Band ordered when downloaded with the package:
1. - 442.7 nm (Coastal aerosol)
2. - 492.4 nm (blue)
3. - 559.8 nm (Green)
4. - 664.6 nm (Red)
5. - 704.1 nm (Vegetation Red Edge)
6. - 740.5 nm (Vegetation Red Edge)
7. - 782.8 nm (Vegetation Red Edge)
8. - 832.8 nm (below 20m rsolution)/864.7 nm (above 20m resolution) (Nir)/ (Vegetation Red Edge)
9. - 945.1 nm (Water vapour)
10. - 1613.7 nm (SWIR)
11. - 2202.4 nm (SWIR)
Let's print RGB pictures just to see.
We will start by adding names to the specific bands
```{r}
pic_1_RGB<-subset(pic_1,c("b2","b3","b4"))
pic_2_RGB<-subset(pic_2,c("b2","b3","b4"))
pic_3_RGB<-subset(pic_2,c("b2","b3","b4"))
```


