---
title: "Proof of concept or illusions? WOrld FOod STudies crop model for Israeli Chickpea"
author: "Omer Perach"
date: '2021-08-24'
output: html_document
slug: []
categories:
- Crop Model
- Spatial Analysis
tags:
- Agriculture
- Modeling
description: ''
topics: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is WOFOST?

WOFOST is a simulation model for the quantitative analysis of the growth and production of annual field crops [Please read](https://www.wur.nl/en/show/A-gentle-introduction-to-WOFOST.htm).  WOFOST could basically help us to understand the dynamic of the plant biological features such as **Leaf area index**, **Biomass**, **Water use**, **Yield** etc. for a location given knowledge about soil, crop, weather and crop management. 

## Application of WOFOST

* Yield risk and inter- annual yield variability.
* Yield variability over soil types, or over a range of agro-hydrological conditions.
* Differences among cultivars.
* Relative importance of growth determining factors.
* Many more

## My plan

WOFOST is a huge and amazing project and I am on the very early stage of learning it. Nonetheless, I believe in learning from doing so I decided to try and take a new R package *Rwofost* and use it. This R package trying to implement the WOFOST model in the R environment. It is worth mentioning that their is a python shell that are more suitable to work with that model [Please read](https://pcse.readthedocs.io/en/stable/).

## What parameters the package functions need to recieve?

The mandatory variables for the functions are:

* crop - which crop I want to model. from a list.
* whether - weather data set with the following parameters:  

  + date-"Date" class variable in R.
  + srad-Solar radiation kJ m^-2 day^-1.
  + tmin-minimum temperature per day (degrees C).
  + tmax-maximum temperature per day (degrees C).
  + vapr-vapor pressure (kPa). 
  + wind-wind speed (m s^-1).
  + prec-precipitation (mm day^-1).

## First stage - single station

In the first stage I will only collect data for a single station. I decided to focus in a data from a station called *Newe Ya'ar*. The station located near a small village called *Ramat Yishai*. The reasons I am choosing to work on this station are:

* I am working on chickpea in my PhD
* The soil type in that reason is clay.
* The area is good representative of a place with annual field crops.
* I have a knowledge on chickpea grown at Gilat  from 2021 season.

## The data from the station

```{r Reading the data set}
df<-read.csv("weather_data.csv")
head(df)
```

## Data preperation

The only state variable which we don't have is vapor pressure (hPa). This can be achieved by using maximum and minimum temperature with relative maximum and minimum humidity [please see equation](http://www.fao.org/3/x0490e/x0490e07.htm). WOFOST unit for all variables:  
* TMAX (Celsius)
* TMIN (Celsius)
* VAP (Vapor Pressure) (Kpa)
* WIND (Wind speed) (meter/second)
* RAIN (cm/day)
* IRRAD (Kjoule/m^2*day)

```{r Computing missing state variables}
df$e0min<-0.6108*exp((17.27*df$TMIN)/(df$TMIN+237.3))
df$e0max<-0.6108*exp((17.27*df$TMAX)/(df$TMAX+237.3))
df$VAP<-((df$e0min*(df$RHMIN/100))+(df$e0max*(df$RHMAX/100)))/2
df$IRRAD<-df$IRRAD*1000
#Handling the date type
library(tidyverse)
df$date_time<-str_remove(df$date_time," ")
df$date_time<-str_remove(df$date_time,"24:00")
df$date_time<-as.Date(df$date_time,format="%d/%m/%Y")

```

This is it, we only need the following 7 variables for the model:  
* TMAX
* TMIN
* VAP
* WIND
* RAIN
* IRRAD

Let's keep only these variables in a new data set

```{r Model running, echo=TRUE}
library(Rwofost)
df_for_model<-df[c("date_time","TMAX","TMIN","VAP","RAIN","WIND","IRRAD")]
##Choosing crop from wofost library
crop<-wofost_crop("chickpea")
##Choosing soil, arbitary (very important but not my area of expertise)
soil<-wofost_soil("ec1")
#Control parameters
contr<-wofost_control()
contr$modelstart<-as.Date("2021-01-12")
contr$latitude<-32.70
contr$elevation<-35
#run model
colnames(df_for_model)
colnames(df_for_model)<-c("date","tmax","tmin","vapr","prec","wind","srad")
d<-wofost(crop,df_for_model,soil,contr)
head(d)
```
```{r echo=TRUE}
p<-d%>%ggplot(aes(x=date,y=WSO))+geom_point()+xlab("Date")+ylab("Grain dry weight Kg/Ha")
print(p)
```


Working, did not check it yet against observed data but the overall message is **we can use WOFOST to model Israeli chickpea**.

To be continued...



  
  

